### 2014-02-14 docs/kkr13/kkr13-eng.tex
kkr

\documentclass[11pt,twoside]{article}
\usepackage{kkr13-eng}
\usepackage{todonotes}
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{booktabs}
\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}

@@

\begin{document}

\listoftodos

@@

  This paper introduces a Bag of Words (BoW) semantic object classification framework based on RGBD data registered by a Microsoft Kinect camera. The BoW concept in natural language processing is introduced and a BoW image processing pipeline is described. Each part of the pipeline (region detection, feature extraction and vector quantization) is characterised and the most successful algorithms are outlined. Applications of BoW in image processing and their results are discussed. Additionally, the problem of classification is described. Algorithms with high performance potential were identified and evaluated. Salient region detection is done by SIFT, ISS3D and Uniform Sampling. Description is performed by FPFH, PFH and PFHRGB. Vector quantization was carried out by the KMeans algorithm. An SVM with the RBF kernel trained with an All-vs-All was used as a classifier. We achieved the accuracy of 65.22\% on 8 categories of the Berkely 3D Object Dataset and 62.3\% on a 10 category dataset compiled by Zhang from the Tokyo University.
}

@@

  It has never been easier to capture visual information. Popular storage services grow lighting fast due to terabytes of photographs and movies uploaded every day. The growth is so fast that hand-tagging and description, the traditional means of annotation, cease to suffice. They are ambiguous, emotional and rarely optimal. With no better solution at hand databases are becoming increasingly harder to browse. Another, but quite similar setting occurs in the field of mobile robotics. A mobile robot is there to interact with its environment. If so, it would be desirable for the robot to know what kind of surroundings it is in. The problem can be treated as a scene or object classification problem, one of the most popular computer vision issues in the last decade. \todo[inline]{ref}

  This paper tackles two issues. Firstly, the Bag of Words (or BoW for short) approach to image classification is reviewed. The Bag of Words method originated from the natural language processing domain \cite{tsai2012bag}. It makes a strong assumption that word occurrence in a text document define its meaning. It were Csurka \textit{et al} who first used BoW for image classification in 2004 \cite{csurka2004visual}. Since images cannot be simply casted into natural language documents a pipeline that help us achieve this goal is needed. It usually consist of: keypoint detection, feature extraction, vector quantization, together with codebook construction, and classification, steps that will be discussed later.

  Being efficient, efective and flexible, a Bag of Words processing pipeline is, however, a blessing in a disguise. While in fact it can provide satisfactory results with little effort, it is cumbersome to tune, thus spectacular performance is hard to achieve. One reason is the abundance of suitable algorithms available. Wide variety of building blocks might make it flexible, but it renders choosing an optimal combination infeasible.

  The rest of this work is organised as follows: Section 2 describes the Bag of Words technique in detail. Construction of an image model is discussed and the most popular algorithms are outlined. Basic information on a problem of classification are introduced the most popular approaches of solving it are discussed. Section 3 describes two renown RGBD image datasets suitable for evaluation of RGBD data-based object classifiers. Section 4 addresses the topics of experimental setup, conducted experiments and provides the results. Finally, Section 5 concludes this work.

@@

  Bag of Words is an intermediate representation in a form of a histogram used in natural language processing, information retrieval and data mining. It obliterates any grammatical information as well as word order in order to perserve word incidence only.

@@

  The figure \ref{fig:bow_example} shows a piece of text and its BoW model, where articles and punctuation marks were left out. In order to compare different documents a global dictionary must be built. The dictionary (codebook) is constructed by taking every word from all available documents and removing duplicates. One can image that if a dictionary is of any considerable size resulting representation of especially small documents will be very sparse. Classification algorithms can be optimised for sparse data in order to boost performance.

@@

  \subsection{Processing Pipeline}	
    Tsai describes a well established pipeline that allows translation of images into text documents \cite{tsai2012bag}. It consists of the following steps: (1) region detection, (2) feature extraction, (3) vector quantization and (4) BoW histogram formation as can be seen in the figure \ref{fig:bow_pipeline}. The author summarises the most commonly used methods for performing these tasks.

@@

  \subsubsection{Region Detection}
    Characteristic region detection is the first step in any Bag of Words framework. Numerous detection methods have been developed, but choosing the right one for any particular case might prove tricky.

    The most common detectors make use of a Harris corner detector or image's first or second derivatives. A Harris-Laplace detector is an example of a Harris-based detector. The Harris function is scale adapted and its outcome is a subject to a Laplacian-of-Gaussian (LoG) operator, which selects relevant points in scale space. Images' regions' 2\textsuperscript{nd} derivatives, namely the regions' Hessians, can be combined with a LoG operator as well. This combination allows selection of points significant in the two spaces: the scale space and the Hessian's determinant space. The latter entails the speed at which pixel intensities change in the neighbourhood of a point.

    A number of more advanced recipes for salient region localisation have been developed and implemented. These include for example SIFT \cite{sift_keypoint}, SUSAN \cite{susan_keypoint} and Intrinistic Shape Signatures \cite{iss_keypoint}. The majority of keypoint detection formulas is being developed for the 2D domain. A number of them have been adapted to 3D, however. A comparative evaluation of detection algorithms available in the PointCloud Library (PCL, discussed below) can be find in \cite{pcl_keypoint_comparision}. Another comprehensive study is \cite{3d_keypoint_eval}. All these formulas, called sparse feature detectors, resort to selection of maxima in specific state spaces. An entirely different scheme is to use a dense feature detector. That is, take a uniformly sampled grid of points. Dense detectors have an advantage in that they sample slow changing regions in terms of gradient or hessian of an image. Examples of such regions would be a clear sky or a calm ocean. Li \emph{et al} showed that dense detectors generally outperform sparse ones \cite{fei2005bayesian}.

  \subsubsection{Feature Extraction}
    Suppose only coordinates of a keypoint were known. In case of even the smallest rotation or translation they would change and the keypoint would be lost. Keypoints should be described in a way that makes them invariant to affine transformations, changes in light intensity or colour saturation. It is hard to achieve all of these properties simultaneously, but methods that meet some of them exist.

    Keypoint description takes form of coordinates in a high-dimensional space. One of the most precise and repeatable algorithm is SIFT \cite{sift_features}, which is a 3D histogram of gradients structured as a 128-dimensional vector of floating point values. It is the most often extracted descriptor in BoW pipelines. Other methods include various colour descriptors, binary descriptors such as 512-dimensional GIST \cite{ponce2011cv}. There are techniques designed for 3D exclusively. Among them are Persistent Point Feature Histogram (PFH) \cite{pfh_rusu2008}, its faster alternative FPFH \cite{fpfh_rusu2009} and PFHRGB, which takes into account colour information, all implemented in PCL.

  \subsubsection{Vector Quantization}
    When features are extracted they have to be normalised. Vector quantization step have two main phases. The first one is a codebook construction from a training dataset. The second phase is responsible for parsing (or translation) raw image descriptors into a form compatible with the newly constructed codebook. The simplest way of building a dictionary is to find patterns or regions within the descriptors computed on a training dataset. Then, the parsing phase is about assigning a descriptor to one of the codebook elements (e.g. to one of the regions). Suppose we use centroids or medoids of the discovered regions as our codebook's entries. Then, we can match a descriptor to one of the centroids performing a nearest neighbour search. The descriptor parsed this way is called a \textit{visual word}. Finally, we build a histogram of visual word occurrence. The histogram have a fixed size equal to the codebook size. Each bin of the histogram tells us how many visual words of a particular kind were in the description of the corresponding image.

    The kMeans is the single most popular vector quantization algorithm used in the BoW pipelines \cite{tsai2012bag}. Developed in 1950's, it is well known and simple to implement. kMeans divides all data into a predefined amount of clusters and computes the clusters' centroids. Many modifications and alternative versions have emerged \cite{kmeans_jain2010data}. Some of them are: faster than the original \emph{approximate kmeans}, \emph{hierarchical kmeans}, which automatically chooses the final number of clusters and a \emph{soft kmeans} --- a variation of the algorithm that allows a fuzzy alignment. The fuzzy alignment mean that each point can belong to several clusters with different weights. A weight can be proportional for instance to the inverse square of the distance to a centroid.

    If computational cost is of no concern, or if required precision is of the utmost priority, a Gaussian Mixture Model (GMM) can be used. The GMM partitions data into a set of clusters, finds their means and covariances. In the parsing step we compute a probability distribution of a descriptor over all the clusters. We can then build a fuzzy aligned histogram taking the probability distribution as weights. The GMM can be thought of as a generalisation of the kMeans algorithm. The drawback of the GMM is its massive computational cost in comparison with still expensive kMeans. Recently, Perronnin \textit{et al} proposed GMM based fisher-kernel vector quantization step with superior results \cite{fisher1, fisher2}.

    After we choose a vector quantization algorithm, we have to decide what size of a dictionary to use. In a simplest case, when the kMeans is used, we can treat each clusters' centroid as a visual word. Therefore, the dictionary size defines how much information we retain in a BoW histogram. A big training set containing many classes with large inter-class and inter-class variance is likely to require a huge codebook. On the other hand, too big a dictionary might introduce quantization artefacts. One has to bear in mind that the vector quantization step is the most time-consuming part of the BoW pipeline. The highest computational complexity is associated with the codebook construction step and it is$O(n^3)$, where $n$ is the codebook size.

    It is possible to combine several feature extraction algorithms before creating the codebook (Early Fusion) or create many codebooks and concatenate resulting histograms (Late Fusion). Both schemes are as simple as vector concatenation. The Early Fusion concatenates two feature vectors if they have been extracted from the same keypoint. The Late Fusion joins histograms, outputted by separate quantization steps.

@@

\section{Problem of Classification}

@@

    \subsubsection{Na\`ive Bayes}
    Assume that we have a \textit{n}-dimensional feature vector $\mathbf{X}$ such that $\mathbf{X} = \left\{x_1, x_2, ... , x_\mathit{n}\right\}$ and an unknown category label \textit{C}. The probability of the category \textit{C} given the feature vector \textbf{X} is $P\left(\mathit{C}\middle|\mathbf{X}\right) = P\left(\mathit{C}\middle|\left\{x_1, x_2, ..., x_\mathit{n}\right\}\right)$. The strong (or na\`ive) Bayes assumption says that features are independent of each other, that is $P(\mathbf{X}) = P(x_1)*P(x_2)*...*P(x_\mathit{n})$. If so, we can compute probability distributions of every class in our training set given every feature. Then, for any new feature vector the joint probability distribution over a set of features can be calculated and an appropriate class label can be assigned.

    \subsubsection{k-Nearest Neighbours (kNN)}
    The nearest neighbours classifier assumes that if there are labelled points in a neighbourhood of a newly added point, then the new point can be classified based on it's neighbours' labels. There is an issue of choosing the correct number of nearest neighbours (or k) to consider.

    \subsubsection{Support Vector Machine (SVM)}
    \begin{figure}[!ht]
    \centering
    \includegraphics[width=0.75\textwidth]{../figs/svm}
    \caption{Classification by a Support Vector Machine. The distance of the separating plane from both datasets is maximal. The figure was originally published in \cite{ponce2011cv}}
    \label{fig:svm}
    \end{figure}

    Assume that a set of pairs $\{\{\mathbf{x}_1, y_1\}, \{\mathbf{x}_2, y_2\}, ...\}$ where $x_\mathit{i}$ are features and $y_\mathit{i} \in \{-1, 1\}$ are labels is given. Further assume that points with different labels are two linearly-separable datasets as shown in the figure \ref{fig:svm}, where dots and circles are points with different labels. Then, parameters \textbf{\textit{w}} and  \textit{b} exist such that	
    \begin{equation}	                        y_\mathit{i}\left(\mathbf{w}*\mathbf{x}_\mathit{i} + b\right) > 0
    \end{equation}
    for every $\mathbf{x}_\mathit{i}$ or a data point. This equation specifies constraints on a plane (or a hyperplane in general) that separates different classes. A Support Vector Machine finds the parameters \textbf{\textit{w}} and  \textit{b} so as to maximise the distance of the plane from members of both classes. We can combine several SVM classifiers in a 1-versus-all or an all-versus-all scheme.

@@

  Datasets are essential for the following reasons: (1) Supervised learning algorithms have ravenous apetite for labelled data. An open database of labelled RGBD images would be of a great help in evaluation of algorithms. (2) If any comparison of algorithms is to be meaningful it should be performed on the same data. There is a multitude of RGB datasets available: ImageNet\footnote{\url{http://www.image-net.org/}}, PascalVoc\footnote{\url{http://pascallin.ecs.soton.ac.uk/challenges/VOC/}}, LabelMe\footnote{\url{http://labelme.csail.mit.edu/Release3.0/browserTools/php/dataset.php}} and SUN\footnote{\url{http://groups.csail.mit.edu/vision/SUN/}}, among others. Unfortunately, they do not contain any depth information and cannot be used for the purpose of object classification based on RGBD data. On the other hand, the vast majority of RGBD datasets are focused on tracking or instance-level recognition, have insufficient number of examples per category or have too few categories. Author managed to find only two datasets suited to his needs. They are: the Berkeley 3D Object (B3DO) dataset \cite{B3DO} and a dataset compiled by Zhang \emph{et al} at the University of Tokyo\cite{zhangcategory}. The latter will be abbreviated Tokyo dataset from now on.

@@

    The Berkeley 3D Object dataset, shown in \refFig{fig:b3do_objects}, was specifically designed for the purpose of object detection and classification. It consists of cluttered images in indoor environment, with many labelled and random objects per image. The images are often poorly lit and objects are partially occluded. There are around 50 classes with more than 20 examples in each of them. RGBD data are provided as pairs of colour images (8 bit RGB jpeg files) and depth maps (16 bit png files). Images are densely labelled --- for every pair of images there is an xml file with annotations. Each annotation states a category of an object as well as it's bounding box location. Since neigher image segmentation nor object detection is addressed in this paper we had to extract objects from the images before the dataset could be used.

@@

\section{ EXPERIMENTS }
  There is a plenty of algorithms suitable for any of the four steps of Bag of Words image processing. Most of them have multiple tunable parameters and their performance might be interdependant on each other. Comprehensive evaluation is inevitable if an optimal combination is to be selected. Considering the PointCloud Library\footnote{\url{http://pointclouds.org/}} (PCL) alone yields far too many options to test, with 12 keypoint detectors and more than 20 feature descriptors. A literature review helped to narrow the scope of evaluation considerably. All experiments were run on a notebook with Intel Core i7 quad core 2.2 GHz CPU, 8 GB of RAM and a nVidia M540 GPU.

  \subsection{Detectors and descriptors}
    Keypoint detection algorithms were compared with respect to repeatability, invariance in \cite{pcl_keypoint_comparision} and in terms of object retrieval and recognition in \cite{3d_keypoint_eval}. 3D-SIFT and ISS both achieve the highest robustness, while the latter appears to be the best choice for object retrieval. Dense detector (or uniform sampling) achieves the best performance in 2D image classification \cite{tsai2012bag}. Therefore, this paper compares SIFT, ISS and Uniform Sampling keypoint detection techniques. Additionaly, PCL descriptors were evaluated in terms of performance in object retrieval in \cite{pcl_features}. PFHRGB and SHOTCOLOR delivered the best results, while their simpler counterparts (PFH, FPFH, SHOT) were only a little worse. USC, which is a global descriptor and thus not suitable for the Bag of Words approach, achieved very similar accuracy. The family of PFH descriptors is evaluated in this paper.

@@

    The best results of each detector/descriptor combination on the B3DO dataset are shown in table \ref{tab:desc_b3do}. ISS consistently proved to be the most successful detection algorithm in this setting. SIFT and Uniform Sampling scored similar accuracy but the latter is significantly faster. SIFT is in fact the slowest of the three. As for the descriptors, FPFH delivered the best performance even though it is an approximate and theoretically the least precise algorithm. It is also the fastest of the three. It appears that the combination of ISS and FPFH is not only surprisingly good in terms of accuracy; it is also the fastest. These two algorithms allow near real time point-cloud processing and classification. PFHRGB and SIFT are several times slower.

@@

  \subsection{B3DO}
    Extracting objects from the B3DO datasets yielded far too many and unbalanced categories. Nnumber of objects in a category varies from 1 (tape) to 299 (table). 8 categories were picked at random with a restriction that there should be at least 50 object instances in a category. Then, the objects were split into train and test set with a proportion of 1:1. They were used for estimation and evaluation respectively.

    The highest accuracy achieved for this dataset is 65.22\%. Table \ref{tab:b3do_conf_matrix} contains a confusion matrix, number of examples per category and accuracy. The confusion matrix depicts classification errors. Let m\subscript{i, j} be an element of the confusion matrix at the i\superscript{th} row and j\superscript{th} column. It shows how many elements from the i\superscript{th} category was assigned to the j\superscript{th} category. A high value of m\subscript{i, j} such that $i \neq j$ indicates that the classifier cannot distinguish those two classes.

@@

    Every of the 343 images from Tokyo dataset were used. Data was split into test and train set in proportions 1:2 in order to provide more training examples due to low number of objects in some classes. Even tough the overall accuracy achieved on the Tokyo dataset is similar in value to that of the B3DO, the structure of the result is very different. It is clearly visible that there is a strong correlation between a per class accuracy and the number of entries in this class. The highest performance in the bicycle category is coupled with the largest number of entries. On the other end of the scale there are cart and printer categories with only 2 and 4 entries respectively. On top of that, differences between some classes are marginal. The majority of carts and baskets were put into the bucket category. It does not surprise, for they are simply akin as can be seen in \refFig{fig:tokyo}.

@@

  In this paper we have tackled the issue of three dimensional point cloud classification with a bag of words-based approach.

  The main focus was put on the design and efficient implementation of the pipeline in a way that allows easy addition of algorithms and switching them at run-time.

  What is more, we have gone to great lengths in order to find scientific databases that are suitable for the task of point cloud classification. We believe that the presence of such databases is fundamental, for it allows meaningful comparison with other researchers. We have experienced the following problems: Firstly, point cloud processing is computationally expensive and inefficient. Due to the lack of structure in point clouds (\textit{i.e.} no predefined relationship between points) additional calculations have to be performed (\textit{e.g.} finding nearest neighbours requires distance comparison for all points in a cloud, whereas for a 2D image one instantly knows where adjacent pixels lie). Secondly, there are virtually no databases that would allow extensive training of the bag of words-based pipeline. Machine learning algorithms have ravenous appetite for labelled data. As the newest findings in the state-of-the-art 2D object classification suggest, an order of tens of thousands of samples is a prerequisite for any satisfactory results We had to make do with as few as 330 objects in our training set. Moreover, there is very little interest in the scientific community for this particular task. The majority of research is conducted on retrieval of CAD-like objects from mesh databases. Last, but not least, there is no 3D detection or description algorithm designed solely for the purpose of bag of words object classification.

  In conclusion I would not recommend any bag-of-words based approach classification of real point clouds registered with a low quality camera such as Microsoft Kinect. Being computationally expensive and rather inefficient, it delivers inferior performance in terms of accuracy and speed when compared to 2D RGB image based algorithms. To makes matters worse, generalisation of 2D algorithms to the 3D domain can be very involved, causing the cutting-edge inventions of the 2D image processing scientists unavailable for the purpose of 3D point cloud processing.

\subsubsection{\small Mathematical formulae}

 The number of a formula has to be written
 in parentheses shifted to the right (see below). Symbols of variables 
 should be written in italic.
 \begin{equation}
   x = \frac{a^{(1-z)} + b}{10 - d_2}
 \end{equation}

 \noindent
 Intervals of values should be written without spaces, e.g.\ 4--45MPa.

\section{FIGURES AND TABLES}

\subsection{Figures and their arrangement}

 Figures should be numbered with the Arabic numerals (see fig. 
 \ref{fig-example}).

 \begin{figure}[hbtp]
  \fbox{\rule{0mm}{45mm}\rule{85mm}{0mm}}
  \caption{The figure caption}
  \label{fig-example}
 \end{figure}

\subsection{Tables}

 Tables have to be formated in the way presented below (see table 
 \ref{tab-example}). The tables have to be numbered  separately from 
 the figures.

>>>>

\documentclass[11pt,twoside]{article}
\usepackage{kkr13-eng}
\usepackage{todonotes}
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{booktabs}
\usepackage{cite}
\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}

@@

\bibliographystyle{abbrv}
\begin{document}

%\listoftodos

@@

  This paper introduces a Bag of Words (BoW) semantic object classification framework based on RGBD data registered by a Microsoft Kinect camera. The BoW concept in natural language processing is introduced and a BoW image processing pipeline is described. Each part of the pipeline (region detection, feature extraction and vector quantization) is characterized and the most successful algorithms are outlined. Applications of BoW in image processing and their results are discussed. Additionally, the problem of classification is described. Algorithms with high performance potential were identified and evaluated. Salient region detection is done by SIFT, ISS3D and Uniform Sampling. Description is performed by FPFH, PFH and PFHRGB. Vector quantization was carried out by the KMeans algorithm. An SVM with the RBF kernel trained with an All-vs-All was used as a classifier. We achieved the accuracy of 65.22\% on 8 categories of the Berkely 3D Object Dataset and 62.3\% on a 10 category dataset compiled by Zhang from the Tokyo University.
}

@@

  It has never been easier to capture visual information. Popular storage services grow lighting fast due to terabytes of photographs and movies uploaded every day. The growth is so fast that hand-tagging and description, the traditional means of annotation, cease to suffice. They are ambiguous, emotional and rarely optimal. With no better solution at hand databases are becoming increasingly harder to browse. Another, but quite similar setting occurs in the field of mobile robotics. A mobile robot is there to interact with its environment. If so, it would be desirable for the robot to know what kind of surroundings it is in. The problem can be treated as a scene or object classification problem, one of the most popular computer vision issues in the last decade \cite{ponce2011cv}.

  This paper tackles two issues. Firstly, the Bag of Words (or BoW for short) approach to image classification is reviewed. The Bag of Words method originated from the natural language processing domain \cite{tsai2012bag}. It makes a strong assumption that word occurrence in a text document define its meaning. It were Csurka \textit{et al} who first used BoW for image classification in 2004 \cite{csurka2004visual}. Since images cannot be simply cast into natural language documents a pipeline that help us achieve this goal is needed. It usually consist of: keypoint detection, feature extraction, vector quantization together with codebook construction and classification, steps that will be discussed later.

  Being efficient, effective and flexible, a Bag of Words processing pipeline is, however, a blessing in a disguise. While in fact it can provide satisfactory results with little effort, it is cumbersome to tune, thus spectacular performance is hard to achieve. One reason is the abundance of suitable algorithms available. Wide variety of building blocks might make it flexible, but it renders choosing an optimal combination infeasible.

  The rest of this work is organised as follows: Section 2 describes the Bag of Words technique in detail. Construction of an image model is discussed and the most popular algorithms are outlined. Then, applications of BoW in computer vision are reported and the problem of classification is introduced. Section 3 describes two renown RGBD image datasets suitable for evaluation of RGBD data-based object classifiers. Section 4 addresses the topics of experimental setup, conducted experiments and provides the results. Finally, Section 5 concludes this work.

@@

  The Bag of Words model is an intermediate representation in a form of a histogram used in natural language processing, information retrieval and data mining. It obliterates any grammatical information as well as word order so as to preserve word incidence only.

@@

  \refFig{fig:bow_example} shows a piece of text and its BoW model, with articles and punctuation marks left out. In order to compare different documents a global dictionary (codebook) must be built. It is constructed by taking every unique word from all available documents. One can image that if a dictionary is of any considerable size the resulting representation of especially small documents can be very sparse. Classification algorithms can be optimised for sparse data in order to boost performance.

@@

  \subsection{Processing Pipeline}	
    Tsai describes a well established pipeline that allows translation of images into text documents \cite{tsai2012bag}. It consists of the following steps: (1) region detection, (2) feature extraction, (3) vector quantization and (4) BoW histogram formation as can be seen in \refFig{fig:bow_pipeline}. The author summarises the most commonly used methods for performing these tasks.

@@

  \subsubsection{Region Detection}
    Characteristic region detection is the first step in any Bag of Words framework. Numerous detection methods have been developed, which might choosing the right one tricky. The most common detectors use a Harris corner detector or 1\textsuperscript{st} or 2\textsuperscript{nd} image derivatives. A Harris-Laplace detector is an example of a Harris-based detector. The Harris function is scale adapted and its outcome is subjected to a Laplacian-of-Gaussian (LoG) operator, which selects points in scale space. 2\textsuperscript{nd} image derivatives~---~Hessians can be combined with a LoG operator in order to further select points in the Hessian's determinant space.

    A number of more advanced recipes for salient region localisation have been developed and implemented. These include for example SIFT \cite{sift_keypoint}, SUSAN \cite{susan_keypoint} and ISS \cite{iss_keypoint}, with the vast majority implemented for the 2D domain and only a few generalised to 3D. A comparative study of detectors available in the PointCloud Library (PCL) can be found in \cite{pcl_keypoint_comparision, 3d_keypoint_eval}. All of the above, called sparse feature detectors, resort to selection of maxima in specific state spaces. An entirely different scheme is to use a dense feature detector, that is to take a uniformly sampled grid of points. Dense detectors are advantageous in that they sample slow changing regions in terms of image derivatives. Examples of such regions are a clear sky or a calm ocean. Li \emph{et al} showed that dense detectors generally outperform sparse ones \cite{fei2005bayesian}.

  \subsubsection{Feature Extraction}
    Suppose only keypoint's coordinates were given. The smallest affine transformation applied to the image would render the keypoints useless. Feature extraction is about describing a region in a way that is invariant to affine transformations, light intensity changes or colour saturation. Achieving all these properties simultaneously can be hard but methods that achieve some of them do exist. Keypoint description takes form of coordinates in a high-dimensional space. One of the most precise and repeatable algorithm is SIFT \cite{sift_features}, which is a 3D histogram of gradients structured as a 128-dimensional vector of floating point values. It is the most often extracted descriptor in BoW pipelines. Other methods include various colour descriptors, binary descriptors such as 512-dimensional GIST \cite{ponce2011cv}. There are techniques designed for 3D exclusively. Among them are Persistent Point Feature Histogram (PFH) \cite{pfh_rusu2008}, its faster alternative FPFH \cite{fpfh_rusu2009} and PFHRGB, which takes into account colour information, all implemented in PCL.

  \subsubsection{Vector Quantization}
    Vector quantization comprises of: (1)~codebook construction from a trainset and (2)~descriptor parsing, performed on a test set only on all data, depending on the setting. The simplest way of building a dictionary is to find patterns or regions within descriptors computed on a trainset and use their centroids or medoids as the codebook's entries. The parsing phase is about assigning a descriptor to one of the codebook elements (e.g. 1-nearest neighbour matching with a centroid). The parsed descriptor is called a \textit{visual word}. Finally, a histogram of visual words can be built. The histogram have a fixed size equal to the codebook size. Each bin of the histogram tells us how many visual words of a particular kind an image contains.

    The kMeans is the single most popular vector quantization algorithm used in the BoW pipelines \cite{tsai2012bag}. Developed in 1950's, it is well known and easy to implement. kMeans divides all data into a predefined amount of clusters and computes the their centroids. Many modifications and alternative versions emerged \cite{kmeans_jain2010data}. Some of them are: faster than the original \emph{approximate kmeans}, \emph{hierarchical kmeans}, which automatically chooses the final number of clusters and a \emph{soft kmeans} --- a variation that allows fuzzy alignment. Fuzzy alignment means that each point can belong to several clusters with different weights (e.g. proportional to the inverse square of a distance to a centroid).

    If computational cost is of no concern or if required precision is of the utmost priority, Gaussian Mixture Model (GMM) can be used. GMM partitions data into a set of clusters, finds their means and covariances. In the parsing step it computes a probability distribution of a descriptor over all the clusters. Fuzzy aligned histogram is built by taking the probability distribution as weights. GMM can be thought of as a generalisation of the kMeans algorithm. One drawback is massive computational cost in comparison with still expensive kMeans. Recently, Perronnin \textit{et al} proposed GMM based fisher-kernel vector quantization step yielding superior results \cite{fisher1, fisher2}.

    After choosing a vector quantization algorithm a codebook size have to be determined. With a standard kMeans implementation the dictionary size defines how much information is retained in a BoW histogram. A big trainset containing many classes with large intra-class and inter-class variance is likely to require a huge codebook. On the other hand, too big a dictionary might introduce quantization artefacts. One has to bear in mind that the vector quantization step is the most time-consuming part of the BoW pipeline. The highest computational complexity is associated with the codebook construction step and it is $O(n^3)$, where $n$ is the codebook size.

    It is possible to combine several feature extraction algorithms before creating the codebook (Early Fusion) or create many codebooks and concatenate resulting histograms (Late Fusion). Both schemes are as simple as vector concatenation. The Early Fusion concatenates two feature vectors if they have been extracted from the same keypoint. The Late Fusion joins histograms that are output of separate quantization steps.

@@

\subsection{Problem of Classification}

@@

@@

  Datasets are essential for the following reasons: (1) Supervised learning algorithms have ravenous appetite for labeled data. An open database of labeled RGBD images would be of a great help in evaluation of algorithms. (2) If any comparison of algorithms is to be meaningful it should be performed on the same data. There is a multitude of RGB datasets available: ImageNet\footnote{\url{http://www.image-net.org/}}, PascalVoc\footnote{\url{http://pascallin.ecs.soton.ac.uk/challenges/VOC/}}, LabelMe\footnote{\url{http://labelme.csail.mit.edu/Release3.0/browserTools/php/dataset.php}} and SUN\footnote{\url{http://groups.csail.mit.edu/vision/SUN/}}, among others. Unfortunately, they do not contain any depth information and cannot be used for the purpose of object classification based on RGBD data. On the other hand, the vast majority of RGBD datasets are focused on tracking or instance-level recognition, have insufficient number of examples per category or have too few categories. Author managed to find only two datasets suited to his needs. They are: the Berkeley 3D Object (B3DO) dataset \cite{B3DO} and a dataset compiled by Zhang \emph{et al} at the University of Tokyo\cite{zhangcategory}. The latter will be abbreviated Tokyo dataset from now on.

@@

    The Berkeley 3D Object dataset, shown in \refFig{fig:b3do_objects}, was specifically designed for the purpose of object detection and classification. It consists of cluttered images in indoor environment, with many labelled and random objects per image. The images are often poorly lit and objects are partially occluded. There are around 50 classes with more than 20 examples in each of them. RGBD data are provided as pairs of colour images (8 bit RGB jpeg files) and depth maps (16 bit png files). Images are densely labelled --- for every pair of images there is an xml file with annotations. Each annotation states a category of an object as well as it's bounding box location. Since neither image segmentation nor object detection is addressed in this paper we had to extract objects from the images before the dataset could be used.

@@

\section{ EXPERIMENTS }
  There is a plenty of algorithms suitable for any of the four steps of Bag of Words image processing. Most of them have multiple tunable parameters and their performance might be interdependent on each other. Comprehensive evaluation is inevitable if an optimal combination is to be selected. Considering the PointCloud Library\footnote{\url{http://pointclouds.org/}} (PCL) alone yields far too many options to test, with 12 keypoint detectors and more than 20 feature descriptors. A literature review helped to narrow the scope of evaluation considerably. All experiments were run on a notebook with Intel Core i7 quad core 2.2 GHz CPU, 8 GB of RAM and a nVidia M540 GPU.

  \subsection{Detectors and descriptors}
    Keypoint detection algorithms were compared with respect to repeatability and invariance in \cite{pcl_keypoint_comparision} as well as in terms of object retrieval and recognition in \cite{3d_keypoint_eval}. 3D-SIFT and ISS both achieve the highest robustness, while the latter appears to be the best choice for object retrieval. Dense detector (or uniform sampling) achieves the best performance in 2D image classification \cite{tsai2012bag}. Therefore, this paper compares SIFT, ISS and Uniform Sampling keypoint detection techniques. Additionally, PCL descriptors were evaluated in terms of object retrieval performance in \cite{pcl_features}. PFHRGB and SHOTCOLOR delivered the best results, while their simpler counterparts (PFH, FPFH, SHOT) were only a little worse. USC, which is a global descriptor and thus not suitable for the Bag of Words approach, achieved very similar accuracy. The family of PFH descriptors is evaluated in this paper.

@@

    The best results of each detector/descriptor combination on the B3DO dataset are shown in table \ref{tab:desc_b3do}. ISS consistently proved to be the most successful detection algorithm in this setting. SIFT and Uniform Sampling scored similar accuracy but the latter is significantly faster. SIFT is in fact the slowest of the three. As for the descriptors, FPFH delivered the best performance, even though it is an approximate and theoretically the least precise algorithm. It is also the fastest of the three. The combination of ISS and FPFH delivers outstanding results at the lowest computational cost. It allows near real time point-cloud processing and classification. PFHRGB and SIFT are several times slower.

@@

  \subsection{B3DO}
    Extracting objects from the B3DO datasets yielded far too many and unbalanced categories. The number of objects in each category varies from 1 (tape) to 299 (table). 8 categories were picked at random with a restriction that there should be at least 50 object instances in a category. Then, the objects were split into train and test set with a proportion of 1:1, which were used for estimation and evaluation respectively.

    The highest accuracy achieved for this dataset is 65.22\%. Table \ref{tab:b3do_conf_matrix} contains a confusion matrix, number of examples per category and accuracy. The confusion matrix depicts classification errors. Let $m_{i, j}$ be an element of the confusion matrix at the $i^{th}$ row and $j^{th}$ column. It shows how many elements from the $i^{th}$ category was assigned to the $j^{th}$ category. A high value of $m_{i, j}$ such that $i \neq j$ indicates that the classifier cannot distinguish those two classes.

@@

    Everyone of the 343 images from the Tokyo dataset were used. Data were split into test and train set in proportions 1:2 in order to provide more training examples due to very limited numbers of objects in some classes. Even tough the overall accuracy achieved on the Tokyo dataset is similar in value to that of the B3DO's, the structure of the result is very different as shown in table \ref{tab:tokyo_conf_matrix}. It is clearly visible that there is a strong correlation between a per class accuracy and the number of entries in this class. The highest performance in the bicycle category is coupled with the largest number of entries. On the other end of the scale there are the printer, scanner and cart categories with 2, 2 and 4 entries respectively and 0\% classification accuracy. On top of that, the differences between some classes are marginal. The majority of carts and baskets were put into the bucket category. It does not surprise, for they are simply akin as can be seen in \refFig{fig:tokyo}.

@@

  In this paper the issue of three dimensional point cloud classification with a bag of words-based approach was tackled. Two scientific databases suitable for the task of point cloud classification were identified and adapted for the purpose of this work. Presence of such datasets is vital --- they act as benchmarks and allow objective comparison of different solutions of the problem. The following problems emerged: (1)~Point cloud processing is computationally expensive and inefficient. Due to the lack of structure in point clouds (\textit{i.e.} no predefined spatial relationships between points) additional calculations have to be performed (\textit{e.g.} finding nearest neighbours requires at least several distance computations). (2)~There are very few and small databases available. Machine learning algorithms have ravenous appetite for labelled data. As the newest findings in the state-of-the-art 2D object classification suggest, an order of tens of thousands of samples is a prerequisite for any satisfactory results. This work had to make do with less than 300 objects in the train sets. (3)~There is very little interest in the scientific community for this particular task. The majority of research is conducted on retrieval of CAD-like objects from mesh databases.

  The achieved results the B3DO and Tokyo datasets are respectively $5.21$ and $6.23$ times higher than random rates ($12.5\%$ and $10\%$). This shows that classification of point clouds is feasible and can be performed in what is close to real time on a mainstream notebook. The accuracy can be further enhanced by generalising 2D algorithms to the 3D domain, while the computational time can be reduced by using a GPU for the most computationally expensive parts of the Bag of Words framework. In conclusion, the depth data does not account for a straightforward improvement in classification accuracy. In fact, it is easier to obtain better results on 2D RGB images, because the best algorithms are not converted to the 3D domain and there is substantially more data available for the 2D.

\bibliography{kkr13}
\end{document}
